Entrainment of low Mach number thermals in stratified domains

### 2019-05-10 tex/stratified_thermals.tex
close to a full rough raft

\received{\today}
\revised{\today}
\accepted{??}%\today}
\submitjournal{ApJ}

@@

\author[0000-0002-3433-4733]{Evan H. Anders}
\affil{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affil{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author[0000-0002-7635-9728]{Daniel Lecoanet}
\affil{Stuff}
\author[0000-0001-8935-219X]{Benjamin P. Brown}
\affil{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affil{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}

@@

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}
Recent observations of solar convection have revealed a convective conundrum.
Power spectra have revealed weaker flows than anticipated at large length scales,
calling into question the existence of so-called ``giant cells'' driven by deep
convection which would manifest as powerful, large-scale motions at the solar
surface. This discrepancy between theory and observations has called into question
our fundamental understanding of convection, sparking fruitful investigations into
the nature of convection in the Sun.

@@

If entropy rain does evolve into downward propagating buoyant vortex rings, it is
important to understand how buoyancy effects the filling factor of these basic
convective elements. In the context of Earth's atmosphere, so-called ``thermals''
are thought to be the nucleus of cloud formation. Thermals are buoyant areas
of fluid which evolve into propagating buoyant vortex rings, and their
evolution in the Boussinesq limit have been well studied in the laboratory
for decades (see e.g., CITE), and more recently have been studied through
Direct Numerical Simulation (DNS) in the laminar and turbulent regime
\citep{lecoanet&jeevanjee2018}. One fundamental result of these studies of
thermals is that they experience a large degree of entrainment: their size expands
with height and their propagation velocity slows despite their buoyant nature.
However, we do not know of a study in which the propagation of these thermals,
and thus the nature of their entrainment, is affected by a 
significant atmospheric stratification.

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{overview_fig.png}
    \caption{
    \label{fig:overview} }
\end{figure}

In the absence of buoyantly-induced entrainment, \citet{brandenburg2016}
suggests that the filling factor, $f$, of vortex rings should decrease like
$f \propto \rho^{\alpha}$ for horizontal compression and $f \propto \rho^{\alpha}$ for 
spherical compression. On the other hand, the filling factor of Boussinesq
thermals \emph{increases} like $f \propto d$, where $d$ is the depth propagated.
In this work, we extend the study of \citet{lecoanet&jeevanjee2018} to study
the propagation of low-Mach number, cold thermals in stratified domains, and how
buoyant entrainment affects the scaling of their filling factor with depth. If
buoyant entrainment is a dominant effect, it is possible that these buoyant vortex rings
would simply grow too large and stall before reaching the bottom of the solar convection zone.
On the other hand, if the compression effects suggested by \citet{brandenburg2016} are the
dominant effect, then it is possible that these thermals could propagate to the bottom of the solar
convection zone, or potentially shrink to a sufficiently small size where thermal dissipation is
significant.

We lay out our work as follows. In section \ref{sec:theory}, we develop a theoretical description
of thermals in a stratified domain. In section \ref{sec:experiment}, we describe and verify
the experiments conducted in this work. In section \ref{sec:results}, we compare the results
of our experiments to the theory developed in section \ref{sec:theory}. Finally, in section
\ref{sec:discussion}, we discuss what our results imply for the entropy rain hypothesis.

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{evolution_colormeshes.png}
    \caption{
    \label{fig:evolution_colormeshes} }
\end{figure}

@@

\subsection{Parameterized description of thermal evolution}
\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{constants.png}
    \caption{
    \label{fig:constants} }
\end{figure}

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{results_panels_d_v_t.png}
    \caption{
    \label{fig:results_d_v_t} }
\end{figure}

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{results_panels_vs_depth.png}
    \caption{
    \label{fig:results_vs_depth} }
\end{figure}

@@

\subsection{Verification of 2D Anelastic approximation}
\begin{figure*}[t!]
    \includegraphics[width=\textwidth]{diff_AN_FC.png}
    \caption{
    \label{fig:diff} }
\end{figure*}

>>>>

\received{\today}
\revised{??}
\accepted{??}%\today}
\submitjournal{ApJ}

@@

\author[0000-0002-3433-4733]{Evan H. Anders}
\affil{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affil{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author[0000-0002-7635-9728]{Daniel Lecoanet}
\affil{Princeton Center for Theoretical Science, Princeton, NJ 08544, USA}
\affil{Department of Astrophysical Sciences, Princeton, NJ 08544, USA}
\author[0000-0001-8935-219X]{Benjamin P. Brown}
\affil{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affil{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}

@@

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}
Recent observations of solar convection have revealed a convective conundrum.
Power spectra have revealed weaker flows than anticipated at large length scales
\citep{hanasoge&all2012, greer&all2015},
calling into question the existence of so-called ``giant cells'' driven by deep
convection which would manifest as powerful, large-scale motions at the solar
surface. This discrepancy between theory and observations has called into question
our fundamental understanding of convection, sparking numerous targeted investigations
the nature of convection in the Sun 
\citep{featherstone&hindman2016, omara&all2016, cossette&rast2016, kapyla&all2017,
hotta2017}.

@@

If entropy rain does evolve into downward propagating buoyant vortex rings, it is
important to understand how buoyancy effects the filling factor of these basic
convective elements. In the context of Earth's atmosphere, so-called ``thermals''
are thought to be the nucleus of cloud formation. Thermals are buoyant areas
of fluid which evolve into propagating buoyant vortex rings, and their
evolution in the Boussinesq limit have been well studied in the laboratory
for decades \citep[see e.g.][]{morton&all1956, scorer1957}, 
and more recently have been studied through
Direct Numerical Simulation (DNS) in the laminar and turbulent regime
\citep{lecoanet&jeevanjee2018}. One fundamental result of these studies of
thermals is that they experience a large degree of entrainment: their size expands
with height and their propagation velocity slows despite their buoyant nature.
However, we do not know of a study in which the propagation of these thermals,
and thus the nature of their entrainment, is affected by a 
significant atmospheric stratification.

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{overview_fig.png}
    \caption{We have plotted the evolution of the filling factor, or radius
	squared, of a buoyant vortex ring with depth (the $n_\rho = 3$ case examined
	later in this work). Overplotted in a thin
	solid line is the prediction for filling factor growth in the Boussinesq
	case, and in the thin dashed line is plotted the prediction for pure horizontal
	compression, as in \citet{brandenburg2016}.
    \label{fig:overview} }
\end{figure}

In the absence of buoyantly-induced entrainment, \citet{brandenburg2016}
suggests that the filling factor, $f$, of vortex rings should decrease like
$f \propto \rho^{-1}$ for horizontal compression and $f \propto \rho^{-2/3}$ for 
spherical compression. On the other hand, the filling factor of Boussinesq
thermals \emph{increases} like $f \propto d^2$, where $d$ is the depth propagated.
These regimes are shown in Fig.~\ref{fig:overview}, and compared to the true
propagation of a thermal in an appreciably stratified environment.

In this work, we extend the study of \citet{lecoanet&jeevanjee2018} to study
the propagation of low-Mach number, cold thermals in stratified domains. We are
specifically interested in how
buoyant entrainment affects the scaling of the thermal radius, or filling factor, with depth. If
buoyant entrainment is a dominant effect, it is possible that entropy rain
would simply grow too large and stall before reaching the bottom of the solar convection zone.
On the other hand, if the compression effects suggested by \citet{brandenburg2016} are the
dominant effect, then it is possible that these thermals could propagate to the bottom of the solar
convection zone, or potentially shrink to a sufficiently small size where thermal dissipation is
significant.

In section \ref{sec:theory}, we develop a theoretical description
of thermals in a stratified domain. In section \ref{sec:experiment}, we describe 
the experiments conducted in this work. In section \ref{sec:results}, we compare the results
of our experiments to the theory developed in section \ref{sec:theory}. Finally, in section
\ref{sec:discussion}, we discuss what our results imply for the entropy rain hypothesis.

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{evolution_colormeshes.png}
    \caption{The evolution of $\rho S_1 r^3$, or the total entropy weighted by
	the cube of the thermal radius, is shown for two thermals. On the left is
	a thermal in a weakly stratified domain with $n_\rho = 1/2$ density scale heights and 
	on the right is a thermal in a strongly stratified domain with $n_\rho = 3$.
	While both start with precisely the same initial condition, the case with low stratification
	expands with depth like the boussinesq case, whereas the strongly stratified thermal
	compresses with depth.
    \label{fig:evolution_colormeshes} }
\end{figure}

@@

\subsection{Parameterized description of thermal evolution}
\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{constants.png}
    \caption{Plotted are time traces for three quantities which we assume to be constant
	in our thermal theory. In (a) and (b), the circulation and buoyancy, divided by their
	maxima, are plotted vs. depth. We see that with increasing stratification, there is marginally
	more detrainment of both of these quantities, but to first order they are constant over the
	evolution of the thermal. In (c) the constant $f = V/r^3$ is plotted; after significant
	noise during the development of the vortex ring, this quantity remains relatively constant, 
	and increasingly approaches a constant with depth.
    \label{fig:constants} }
\end{figure}

@@

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{results_panels_d_v_t.png}
    \caption{Shown are the measured depths of thermals as a function of time for all 2D
	Anelastic and 3D Fully compressible simulations conducted in this work. Overplotted
	is the theoretical prediction for depth as a function of time.
    \label{fig:results_d_v_t} }
\end{figure}

\begin{figure}[t!]
    \includegraphics[width=\columnwidth]{results_panels_vs_depth.png}
    \caption{Shown are (a) the measured radii of thermals as a function of depth and
	(b) the measured thermal velocities as a function of depth for all 2D
	Anelastic and 3D Fully compressible simulations conducted in this work. Overplotted
	are the predictions from theory.
    \label{fig:results_vs_depth} }
\end{figure}

@@

\subsection{Verification of 2D Anelastic approximation}
\begin{figure*}[t!]
    \includegraphics[width=\textwidth]{diff_AN_FC.png}
    \caption{Measured values of (a) depth vs. time and (c) radius vs. depth are plotted
	for both the 2D anelastic and 3D fully compressible simulations. The fractional
	difference between anelastic and fully compressible results are respectively shown
	in (b) and (d).
    \label{fig:diff} }
\end{figure*}
